### 2013-05-31 doc/nips2013/discussion.tex
spell check, mods to experiments

\section{Related work}

% Observable operator models
HMMs \cite{hsu09spectral},
low-treewidth graphs \cite{parikh12spectral},
probabilistic context-free grammars \cite{cohen12pcfg},
weighted finite state automata \cite{balle12automata},
finite state transducers \cite{balle11transducer}.

% Method of moments
Latent Dirichlet Allocation \cite{anandkumar12lda},
mixture of spherical Gaussians \cite{hsu13spherical},
latent parse tree models \cite{hsu12identifiability}
mixture of linear regressions \cite{chaganty13regression}.
linear Bayesian networks \cite{anandkumar12linear}.

% Structure recovery
structure recovery
\cite{anandkumar11tree}

% Alternate estimators
There are multiple consistent estimators.
Composite likelihood \cite{besag75pseudo,lindsay88composite,liang08asymptotics}
We are assuming inference is tractable.
Our main problem is non-convexity.

\section{Conclusion}

That's it.

>>>>

\section{Discussion}

In recent years, there has been much interest
on using the method of moments and spectral decomposition
for learning latent-variable models.
One line of work uses observable operator models to learn
HMMs \cite{hsu09spectral},
low-treewidth graphs \cite{parikh12spectral},
probabilistic context-free grammars \cite{cohen12pcfg},
weighted finite state automata \cite{balle12automata},
finite state transducers \cite{balle11transducer},
but do not produce parameter estimates.

% Method of moments
Our work builds on a second line of work which focuses
on parameter estimation.  These methods have been used to learn
hidden Markov models \cite{anandkumar12moments},
Latent Dirichlet Allocation \cite{anandkumar12lda},
linear Bayesian networks \cite{anandkumar12linear},
mixture of spherical Gaussians \cite{hsu13spherical},
latent parse tree models \cite{hsu12identifiability}, and
mixture of linear regressions \cite{chaganty13regression}.
Not only can actual parameter estimates be valuable for scientific
applications, they also provide a natural initialization point
for further local optimization, which has been found to improve accuracy
\cite{chaganty13regression}.

Extending this latter perspective,
our work departs from the standard goal of
insisting on constructing consistent parameter estimators, but rather strives
to use spectral decomposition to obtain constraints to improve
parameter estimation.  Using these techniques,
we found empirically that our approach provides
improvements in various models such as the grid model and factorial HMM
for which we do not have a consistent parameter estimate.
We believe using partial information from factorized moments
in conjunction with likelihood optimization
broadens the toolbox of techniques for learning
latent-variable models.
Finally, our technique still relies fundamentally on
having three-view bottlenecks \cite{anandkumar12moments}.
How to extract partial information without this assumption
remains an interesting open problem.

% Structure recovery
%structure recovery
%\cite{anandkumar11tree}

% Alternate estimators
%There are multiple consistent estimators.
%Composite likelihood \cite{besag75pseudo,lindsay88composite,liang08asymptotics}

### 2013-05-31 doc/nips2013/measurements.tex
spell check, mods to experiments

% Examples
\reffig{generalModels} shows examples of log-linear models.
For notational simplicitly, assume each latent variable $h_j \in [\nh]$ is an integer
and each observed variable $x_j \in \R^\ell$ is an indicator vector encoding
one of $\ell$
possible values.

@@

\paragraph{(b) Grid model}
Analogous to the second HMM reduction, let $h_{ij}$ be the bottleneck,
where $(i,j)$ is chosen randomly.
Let $x_{ij}^a,x_{ij}^b$ be two of the views, and any other observed variable as the third view.
$\TensorFactorize$ would reveal conditional means $\E[x_{ij}^a \mid h_{ij}]$
and $\E[x_{ij}^b \mid h_{ij}]$,
and a third quantity that involves the latent-to-latent features and the observation features.
This quantity involves marginalizing out the rest of the graph,
and therefore difficult to use.
%to extract something useful like $\BP[h_{ij}, h_{i(j+1)}]$
%in closed form.
%Note that the analog of this was still possible in the first HMM reduction that
%used only the first three observations, but not in the second one.

@@

In each of these three models, we could extract marginals $\pi = \BP[h_j] \in \R^{\nh}$
and some conditional mean matrix $B = \E[x_{j'} \mid h_j] \in \R^{\nphix \times \nh}$,
but these are insufficient to identify the parameters.
How can we still use this partial information $(\pi,B)$?
To this end, we adopt the measurements framework of \citet{liang09measurements},
which allows us to incorporate moment constraints into a likelihood-based objective in a coherent way.
The basic idea is to define two families of distributions:
$\sQ$, which are consistent with all information; and
$\sP$, the exponential family.  The
framework tries to choose $q \in \sQ$ and $p \in \sP$ to minimize
$\KL{q}{p}$.

>>>>

% Examples
\reffig{generalModels} shows examples of log-linear models.
For notational simplicity, assume each latent variable $h_j \in [\nh]$ is an integer
and each observed variable $x_j \in \R^\ell$ is an indicator vector encoding
one of $\ell$
possible values.

@@

\paragraph{(b) Grid model}
Analogous to the second HMM reduction, let $h_{ij}$ be the bottleneck,
where $(i,j)$ is chosen randomly.
Let $x_{ij}^a,x_{ij}^b$ be two of the views, and any other observed variable as the third view.
$\TensorFactorize$ would reveal conditional means $\E[x_{ij}^a \mid h_{ij}]$
and $\E[x_{ij}^b \mid h_{ij}]$,
and a third quantity that involves the latent-to-latent features and the observation features.
This quantity involves marginalizing out the rest of the graph,
and therefore difficult to use.
Note that if each latent variable had only one observation,
there would be no bottlenecks, and our method would not be applicable.
%to extract something useful like $\BP[h_{ij}, h_{i(j+1)}]$
%in closed form.
%Note that the analog of this was still possible in the first HMM reduction that
%used only the first three observations, but not in the second one.

@@

In each of these three models, we could extract marginals $\pi = \BP[h_j] \in \R^{\nh}$
and some conditional mean matrix $B = \E[x_{j'} \mid h_j] \in \R^{\nphix \times \nh}$,
but these are insufficient to identify the parameters.
How can we still use this partial information $(\pi,B)$?
To this end, we adopt the measurements framework of \citet{liang09measurements}
(also, see \cite{graca08em}),
which allows us to incorporate moment constraints into a likelihood-based objective in a coherent way.
The basic idea is to define two families of distributions:
$\sQ$, which are consistent with all information; and
$\sP$, the exponential family.  The
framework tries to choose $q \in \sQ$ and $p \in \sP$ to minimize
$\KL{q}{p}$.

### 2013-05-31 doc/nips2013/mixture.tex
spell check, mods to experiments

Define
$\pi \eqdef (\BP[h=1], \dots, \BP[h=\nh])$ to be the marginal probabilities
over $h$,
and define the matrix $B \eqdef (\E[\phix(x_j) \mid h])_{h \in [\nh]} \in \R^{\ell \times k}$,
where each column is the expected transformated data conditioned on $h$.

@@

We thus obtain a consistent estimator for $\theta$ assuming the exponential
family is identifiable from fully-observed data (that is
the log-artition function $A(\theta)$ is strongly convex).
%\begin{assumption}
%The log-linear model is identifiable from fully-observed data.
%Equivalently, the log-partition function $A(\theta)$ is strongly convex.
%\end{assumption}
%Identifiability means that the moments $\mu$ uniquely determine $\theta^*$.
%By the moment-generating properties of the log-partition function,
%the optimality condition is that $\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$.
In other words,
tensor factorization (Step 2) yields the mean parameters of the exponential family
($\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$),
which can be mapped to (canonical) parameters $\theta$ via convex optimization.

>>>>

Define
$\pi \eqdef (\BP[h=1], \dots, \BP[h=\nh])$ to be the marginal probabilities
over $h$,
and define the matrix $B \eqdef (\E[\phix(x_j) \mid h])_{h \in [\nh]} \in \R^{\ell \times k}$,
where each column is the expected transformed data conditioned on $h$.

@@

We thus obtain a consistent estimator for $\theta$ assuming the exponential
family is identifiable from fully-observed data (that is
the log-partition function $A(\theta)$ is strongly convex).
%\begin{assumption}
%The log-linear model is identifiable from fully-observed data.
%Equivalently, the log-partition function $A(\theta)$ is strongly convex.
%\end{assumption}
%Identifiability means that the moments $\mu$ uniquely determine $\theta^*$.
%By the moment-generating properties of the log-partition function,
%the optimality condition is that $\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$.
In other words,
tensor factorization (Step 2) yields the mean parameters of the exponential family
($\mu = \nabla A(\theta) = \E_\theta[\phi(x, h)]$),
which can be mapped to (canonical) parameters $\theta$ via convex optimization.
